---
title: "Data Visualisation with Cars Project"
output:

---

###Data Visualisation with Cars Project

This project will use data visualisation techniques to analyse statistics in for cars. 

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models). 

These columns include mpg representing miles/(US) gallon, cyl showing the	number of cylinders
and disp representing	displacement.	hp is the gross horsepower, drat	shows the rear axle ratio
and wt represents	weight in 1000 lbs. qsec is	1/4 mile time, vs	is an engine with 0 = V-shaped, 1 = straight.	am shows the transmission with 0 = automatic and 1 = manual. The number of forward gears is shown as gear with carb as the	number of carburetors.

##Explore the Data

First, the mtcars data set will be loaded from R and the first fives lines of data can be viewed below.

```{r}
#load datasets
library(datasets)

#load mtcars
data("mtcars")

#view first 5 lines
head(mtcars)

```

This data frame with 32 observations on 11 numerical variables.

##Automatic Cars

This project will put a focus on automatic cars. Therefore, a frequency count will be created with the automatic(am) column and reflected in a pie chart.
```{r}

#create table of the automatic column
counts <- table(mtcars$am)

#view pie chart
pie(counts)
```
We can see that the as 0 represents automatics and 1 represents manuals that the majority of cars in the data set are automatic.

##Correlation
To understand the data set further a correlation plot will be constructed using all the variables. 

```{r}
library(ggplot2)
library(corrplot)
corrplot(cor(mtcars),
          type="lower", 
          method="color", 
         addCoef.col="black", 
         diag=FALSE, 
         tl.col="black",
         tl.srt=45, 
         col=heat.colors(10))
```

From the correlation plot, there is a strong positive correlation between automatic cars and miles per gallon(mpg) and rear axle ratio (drat.) With a negative correlation between automatics and weight by 1000 lbs(wt.)This suggests that automatic cars do have influence the miles per gallon and the rear axle ratio. 

This is what we will investigate further. 

##Boxplots

The miles per gallon of a car will be shown below. 

```{r}
#load the packages
library(ggplot2)
library(gridExtra)
library(tidyverse)

#create the boxplot
ggplot(mtcars, aes(x=am, y=mpg, color="age_group"))+
geom_boxplot() +
ggtitle("Miles per Gallon vs. Automatic")

```
We can see that the mean and average miles per hour in the data set is 19. The interquartile range is between 15 and 22.




